---
title: "NMF_Visualization"
author: "Candice Yu"
date: "2023-02-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages 
```{r}
#install.packages("plotly")
library(Biobase)
library(plotly)
library(NMF)
library(tidyverse)
library("plyr")
library("dplyr")
```

# Import data 
```{r}
load('/Users/candiceyu/Documents/PSTAT296B/Code/nmfResults296_20230216.RData')
snmfl$fit$`15`

basismap(snmfl$fit$`15`)
coefmap(minfit(snmfl$fit$`15`))
#coefmap(snmfl$fit$`15`)

```

# Extract the index of the features according to thresholds 
```{r}
# Threshold greater than 0.8
Features_Num_0.8 <- extractFeatures(snmfl$fit$`15`, 0.8)

# Threshold greater between 0.6 to 0.8
Features_Num_0.6 <- extractFeatures(snmfl$fit$`15`, 0.6)
Features_Num_0.6_0.8 <- Features_Num_0.6
for (i in 1:15){
  SameIndex <- !(Features_Num_0.6[[i]]  %in% Features_Num_0.8[[i]])
  Features_Num_0.6_0.8[[i]]<- Features_Num_0.6[[i]][SameIndex]
}

# Threshold greater between 0.4 to 0.6
Features_Num_0.4 <- extractFeatures(snmfl$fit$`15`, 0.4)
Features_Num_0.4_0.6 <- Features_Num_0.4
for (i in 1:15){
  SameIndex <- !(Features_Num_0.4[[i]]  %in% Features_Num_0.6[[i]])
  Features_Num_0.4_0.6[[i]]<- Features_Num_0.4[[i]][SameIndex]
}

# Threshold greater between 0.2 to 0.4
Features_Num_0.2 <- extractFeatures(snmfl$fit$`15`, 0.2)
Features_Num_0.2_0.4 <- Features_Num_0.2
for (i in 1:15){
  SameIndex <- !(Features_Num_0.2[[i]]  %in% Features_Num_0.4[[i]])
  Features_Num_0.2_0.4[[i]]<- Features_Num_0.2[[i]][SameIndex]
}

# Threshold greater between 0 to 0.2
Features_Num_0.0 <- extractFeatures(snmfl$fit$`15`, 0.0)
Features_Num_0.0_0.2 <- Features_Num_0.0
for (i in 1:15){
  SameIndex <- !(Features_Num_0.0[[i]]  %in% Features_Num_0.2[[i]])
  Features_Num_0.0_0.2[[i]]<- Features_Num_0.0[[i]][SameIndex]
}
```

# Extract coef data from final NMF model 
```{r}
basis_df <- data.frame(basis(snmfl$fit$`15`))
basis_df <- cbind(rownames(basis_df), basis_df)
rownames(basis_df) <- NULL
colnames(basis_df)<- c("Benefit","Basis_1", "Basis_2", "Basis_3", "Basis_4", "Basis_5", 
                       "Basis_6", "Basis_7", "Basis_8", "Basis_9", "Basis_10", "Basis_11", 
                       "Basis_12", "Basis_13" ,"Basis_14", "Basis_15")
```

# Extract Benefit for each threshold 
```{r}
# Create Threshold greater than 0.8
Features_0.8 <- combine(basis_df[Features_Num_0.8[1][[1]], c(1,2)], 
                        basis_df[Features_Num_0.8[2][[1]], c(1,3)], 
                        basis_df[Features_Num_0.8[3][[1]], c(1,4)], 
                        basis_df[Features_Num_0.8[4][[1]], c(1,5)], 
                        basis_df[Features_Num_0.8[5][[1]], c(1,6)], 
                        basis_df[Features_Num_0.8[6][[1]], c(1,7)], 
                        basis_df[Features_Num_0.8[7][[1]], c(1,8)], 
                        basis_df[Features_Num_0.8[8][[1]], c(1,9)], 
                        basis_df[Features_Num_0.8[9][[1]], c(1,10)], 
                        basis_df[Features_Num_0.8[10][[1]], c(1,11)], 
                        basis_df[Features_Num_0.8[11][[1]], c(1,12)], 
                        basis_df[Features_Num_0.8[12][[1]], c(1,13)], 
                        basis_df[Features_Num_0.8[13][[1]], c(1,14)], 
                        basis_df[Features_Num_0.8[14][[1]], c(1,15)], 
                        basis_df[Features_Num_0.8[15][[1]], c(1,16)])
Features_0.8_long <- Features_0.8 %>% 
  pivot_longer(cols = 2:16)
Features_0.8_long <- Features_0.8_long[!(is.na(Features_0.8_long$value)), ]
Features_0.8_long$Threshold <- "0.8"

# Threshold greater between 0.6 to 0.8
Features_0.6_0.8 <- combine(basis_df[Features_Num_0.6_0.8[1][[1]], c(1,2)], 
                        basis_df[Features_Num_0.6_0.8[2][[1]], c(1,3)], 
                        basis_df[Features_Num_0.6_0.8[3][[1]], c(1,4)], 
                        basis_df[Features_Num_0.6_0.8[4][[1]], c(1,5)],
                        basis_df[Features_Num_0.6_0.8[5][[1]], c(1,6)], 
                        basis_df[Features_Num_0.6_0.8[6][[1]], c(1,7)], 
                        basis_df[Features_Num_0.6_0.8[7][[1]], c(1,8)],
                        basis_df[Features_Num_0.6_0.8[8][[1]], c(1,9)], 
                        basis_df[Features_Num_0.6_0.8[9][[1]], c(1,10)], 
                        basis_df[Features_Num_0.6_0.8[10][[1]], c(1,11)], 
                        basis_df[Features_Num_0.6_0.8[11][[1]], c(1,12)], 
                        basis_df[Features_Num_0.6_0.8[12][[1]], c(1,13)], 
                        basis_df[Features_Num_0.6_0.8[13][[1]], c(1,14)], 
                        basis_df[Features_Num_0.6_0.8[14][[1]], c(1,15)], 
                        basis_df[Features_Num_0.6_0.8[15][[1]], c(1,16)])
                        
Features_0.6_0.8_long <- Features_0.6_0.8 %>% 
  pivot_longer(cols = 2:16)
Features_0.6_0.8_long <- Features_0.6_0.8_long[!(is.na(Features_0.6_0.8_long$value)), ]
Features_0.6_0.8_long$Threshold <- "0.6_0.8"

# Threshold greater between 0.4 to 0.6
Features_0.4_0.6 <- combine(basis_df[Features_Num_0.4_0.6[1][[1]], c(1,2)], 
                        basis_df[Features_Num_0.4_0.6[2][[1]], c(1,3)], 
                        basis_df[Features_Num_0.4_0.6[3][[1]], c(1,4)], 
                        basis_df[Features_Num_0.4_0.6[4][[1]], c(1,5)],
                        basis_df[Features_Num_0.4_0.6[5][[1]], c(1,6)], 
                        basis_df[Features_Num_0.4_0.6[6][[1]], c(1,7)], 
                        basis_df[Features_Num_0.4_0.6[7][[1]], c(1,8)], 
                        basis_df[Features_Num_0.4_0.6[8][[1]], c(1,9)], 
                        basis_df[Features_Num_0.4_0.6[9][[1]], c(1,10)], 
                        basis_df[Features_Num_0.4_0.6[10][[1]], c(1,11)], 
                        basis_df[Features_Num_0.4_0.6[11][[1]], c(1,12)], 
                        basis_df[Features_Num_0.4_0.6[12][[1]], c(1,13)], 
                        basis_df[Features_Num_0.4_0.6[13][[1]], c(1,14)], 
                        basis_df[Features_Num_0.4_0.6[14][[1]], c(1,15)], 
                        basis_df[Features_Num_0.4_0.6[15][[1]], c(1,16)])
Features_0.4_0.6_long <- Features_0.4_0.6 %>% 
  pivot_longer(cols = 2:16)
Features_0.4_0.6_long <- Features_0.4_0.6_long[!(is.na(Features_0.4_0.6_long$value)), ]
Features_0.4_0.6_long$Threshold <- "0.4_0.6"

# Threshold greater between 0.2 to 0.4
Features_0.2_0.4 <- combine(basis_df[Features_Num_0.2_0.4[1][[1]], c(1,2)], 
                        basis_df[Features_Num_0.2_0.4[2][[1]], c(1,3)], 
                        basis_df[Features_Num_0.2_0.4[3][[1]], c(1,4)], 
                        basis_df[Features_Num_0.2_0.4[4][[1]], c(1,5)],
                        basis_df[Features_Num_0.2_0.4[5][[1]], c(1,6)], 
                        basis_df[Features_Num_0.2_0.4[6][[1]], c(1,7)], 
                        basis_df[Features_Num_0.2_0.4[7][[1]], c(1,8)],
                        basis_df[Features_Num_0.2_0.4[8][[1]], c(1,9)], 
                        basis_df[Features_Num_0.2_0.4[9][[1]], c(1,10)], 
                        basis_df[Features_Num_0.2_0.4[10][[1]], c(1,11)], 
                        basis_df[Features_Num_0.2_0.4[11][[1]], c(1,12)], 
                        basis_df[Features_Num_0.2_0.4[12][[1]], c(1,13)], 
                        basis_df[Features_Num_0.2_0.4[13][[1]], c(1,14)], 
                        basis_df[Features_Num_0.2_0.4[14][[1]], c(1,15)], 
                        basis_df[Features_Num_0.2_0.4[15][[1]], c(1,16)])
Features_0.2_0.4_long <- Features_0.2_0.4 %>% 
  pivot_longer(cols = 2:16)
Features_0.2_0.4_long <- Features_0.2_0.4_long[!(is.na(Features_0.2_0.4_long$value)), ]
Features_0.2_0.4_long$Threshold <- "0.2_0.4"

# Threshold greater between 0.0 to 0.2
Features_0.0_0.2 <- combine(basis_df[Features_Num_0.0_0.2[1][[1]], c(1,2)], 
                        basis_df[Features_Num_0.0_0.2[2][[1]], c(1,3)], 
                        basis_df[Features_Num_0.0_0.2[3][[1]], c(1,4)], 
                        basis_df[Features_Num_0.0_0.2[4][[1]], c(1,5)],
                        basis_df[Features_Num_0.0_0.2[5][[1]], c(1,6)], 
                        basis_df[Features_Num_0.0_0.2[6][[1]], c(1,7)], 
                        basis_df[Features_Num_0.0_0.2[7][[1]], c(1,8)],
                        basis_df[Features_Num_0.0_0.2[8][[1]], c(1,9)], 
                        basis_df[Features_Num_0.0_0.2[9][[1]], c(1,10)], 
                        basis_df[Features_Num_0.0_0.2[10][[1]], c(1,11)], 
                        basis_df[Features_Num_0.0_0.2[11][[1]], c(1,12)], 
                        basis_df[Features_Num_0.0_0.2[12][[1]], c(1,13)], 
                        basis_df[Features_Num_0.0_0.2[13][[1]], c(1,14)], 
                        basis_df[Features_Num_0.0_0.2[14][[1]], c(1,15)], 
                        basis_df[Features_Num_0.0_0.2[15][[1]], c(1,16)])
Features_0.0_0.2_long <- Features_0.0_0.2 %>% 
  pivot_longer(cols = 2:16)
Features_0.0_0.2_long <- Features_0.0_0.2_long[!(is.na(Features_0.0_0.2_long$value)), ]
Features_0.0_0.2_long$Threshold <- "0.0_0.2"

# Combine all threshold into one dataset 
Features_threshold_all <- rbind(Features_0.8_long, Features_0.6_0.8_long, Features_0.4_0.6_long, 
                                Features_0.2_0.4_long, Features_0.0_0.2_long)
names(Features_threshold_all) <- c("Benefit", "Basis", "Coef", "Threshold")
```

# Data frame containing the length of the benefit for each basis
```{r}
length_rank_0.8 <- data.frame()
length_rank_0.6_0.8 <- data.frame()
length_rank_0.4_0.6 <- data.frame()
length_rank_0.2_0.4 <- data.frame()
length_rank_0.0_0.2 <- data.frame()

for (i in 1:7){
  length_rank_0.8 <- rbind(length_rank_0.8, length(Features_Num_0.8[[i]]))
}
for (i in 1:7){
  length_rank_0.6_0.8 <- rbind(length_rank_0.6_0.8, length(Features_Num_0.6_0.8[[i]]))
}
for (i in 1:7){
  length_rank_0.4_0.6 <- rbind(length_rank_0.4_0.6, length(Features_Num_0.4_0.6[[i]]))
}
for (i in 1:7){
  length_rank_0.2_0.4 <- rbind(length_rank_0.2_0.4, length(Features_Num_0.2_0.4[[i]]))
}
for (i in 1:7){
  length_rank_0.0_0.2 <- rbind(length_rank_0.0_0.2, length(Features_Num_0.0_0.2[[i]]))
}
length_rank <- cbind(length_rank_0.8, length_rank_0.6_0.8, length_rank_0.4_0.6, 
                     length_rank_0.2_0.4, length_rank_0.0_0.2)


length_rank <- cbind(rownames(length_rank), length_rank)
rownames(length_rank) <- NULL
colnames(length_rank)<- c("Rank","Length_Rank_0.8", "Length_Rank_0.6_0.8", "Length_Rank_0.4_0.6", 
                       "Length_Rank_0.2_0.4", "Length_Rank_0.0_0.2")
```


# Change the length_rank data to longer version 
```{r}
length_rank_t <- data.frame(t(length_rank))
length_rank_t <- length_rank_t[-1,]

length_rank_t <- cbind(rownames(length_rank_t), length_rank_t)
rownames(length_rank_t) <- NULL
colnames(length_rank_t)<- c("Threshold","Basis_1", "Basis_2", "Basis_3", "Basis_4", "Basis_5", "Basis_6", "Basis_7")

length_rank_long <- length_rank_t %>% 
  pivot_longer(cols = 2:8)
colnames(length_rank_long)<- c("Threshold", "Basis", "Count")
length_rank_long$Count <- as.integer(length_rank_long$Count)
```

# Create a new dataset (1. all benefit separated by group; 2. add column of count)
```{r eval=FALSE}
Features_threshold_combine<- Features_threshold_all %>%
  arrange(desc(Coef)) %>%
  group_by(Basis, Threshold) %>%
  mutate(count=n()) %>%
  mutate(Benefits_all= paste(Benefit, collapse = ", ")) %>%
  subset(select = -c(1,3)) %>%
  unique()


Features_threshold_final <- cbind(Features_threshold_combine, 
      str_split_fixed(Features_threshold_combine$Benefits_all, ", ", n=max(Features_threshold_combine$count)))
```

# Stacked Bar Plot: Benefits in Each Basis Factor Ordered by Threshold
```{r eval=FALSE}
index <- list(names(Features_threshold_final)[5:(max(Features_threshold_combine$count)+4)])
plot_ly(Features_threshold_final, 
        x = ~Basis,
        y = ~count,
        color = ~Threshold,
        type = "bar", 
        hovertext = ~paste(...5, ...6, ...7, ...8, ...9, ...10, ...11, ...12, ...13, ...14, ...15, ...16, ...17,
                      ...18, ...19, ...20, ...21, ...22, ...23, ...24, ...25, ...26, ...27, ...28, ...29, ...30,
                      ...31, ...32, ...33, ...34, ...35, ...36, ...37, ...38, ...39, ...40, ...41, ...42, ...43,
                      ...44, ...45, ...46, ...47, ...48, ...49, ...50, ...51, ...52, ...53, ...54, ...55, ...56,
                      ...57, ...58, ...59, ...60, ...61, ...62, ...63, ...64, ...65, ...66, ...67, ...68, ...69,
                      ...70, ...71, ...72, ...73, ...74, ...75, ...76, ...77, ...78, ...79, ...80, ...81, ...82,
                      ...83, ...84, ...85, ...86, ...87, ...88, ...89, ...90, ...91, ...92, ...93, ...94, ...95,
                      ...96, ...97, ...98, ...99, ...100, ...101, ...102, ...103, ...104, ...105, sep = "\n")) %>%
  layout(yaxis = list(title = 'Count of Benefits'), barmode = "stack",
         title="Benefits in Each Basis Factor Ordered by Threshold",
         hoverlabel = list(font = list(family = "Sitka Small", size = 12, color = "Black"))) 
```


```{r}
# Combine all threshold into one dataset 
Features_threshold_all_2 <- rbind(Features_0.8_long, Features_0.6_0.8_long, Features_0.4_0.6_long, 
                                Features_0.2_0.4_long)
names(Features_threshold_all_2) <- c("Benefit", "Basis", "Coef", "Threshold")
# Create a new dataset (1. all benefit separated by group; 2. add column of count)

Features_threshold_combine_2 <- Features_threshold_all_2 %>%
  arrange(desc(Coef)) %>%
  group_by(Basis, Threshold) %>%
  dplyr::mutate(count=n()) %>%
  dplyr::mutate(Benefits_all= paste(Benefit, collapse = ", ")) %>%
  subset(select = -c(1,3)) %>%
  unique()

Features_threshold_final_2 <- cbind(Features_threshold_combine_2, 
      str_split_fixed(Features_threshold_combine_2$Benefits_all, ", ", n=max(Features_threshold_combine_2$count)))

```

# Stacked Bar Plot: Benefits in Each Basis Factor Ordered by Threshold
```{r}
index <- list(names(Features_threshold_final_2)[5:(max(Features_threshold_combine_2$count)+4)])
plot_ly(Features_threshold_final_2, 
        x = ~Basis,
        y = ~count,
        color = ~Threshold,
#        colors = 'Blues',
        type = "bar", 
        hovertext = ~paste(...5, ...6, ...7, ...8, ...9, ...10, ...11, ...12, ...13, ...14, ...15, ...16, ...17,
                           sep = "\n")) %>%
  layout(xaxis = list(title = "Basis Components", categoryorder = "array", 
                      categoryarray = c("Basis_1", "Basis_2", "Basis_3", "Basis_4", "Basis_5",
                                        "Basis_6", "Basis_7", "Basis_8", "Basis_9", "Basis_10",
                                        "Basis_11", "Basis_12", "Basis_13", "Basis_14", "Basis_15")),
         yaxis = list(title = 'Count of Benefits'), barmode = "stack",
         title="Benefits in Each Basis Component Ordered by Threshold",
         hoverlabel = list(font = list(family = "Sitka Small", size = 12, color = "black"))) 

```

```{r}

Data_Dictionary<- read.csv("/Users/candiceyu/Documents/PSTAT296B/Data/Data Dictionary - Sheet1.csv", header = FALSE)
names(Data_Dictionary) <- c("Benefit","Dictionary")


# Left join on multiple columns
Features_Threshold_Dictionary <- merge(x = Features_threshold_all_2, y = Data_Dictionary, 
                         by = c("Benefit"), all.x=TRUE)




write_csv(Features_Threshold_Dictionary,
          '/Users/candiceyu/Documents/PSTAT296B/Data/Features_Threshold_Dictionary_20230222.csv')

```

